Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2017-12-23 paper-ieee-trans.tex
Update paper-ieee-trans.tex

\IEEEPARstart{W}hen e-Infrastructures in Europe were first conceived, new
services to access distributed resources -- data, computing, instrumental --
were initially considered. Major investments were then injected in order to
develop distributed solutions that provided a seamless and transparent access
to the resources, available for the researchers. One clear example is Worldwide
Large Hadron Collider Computing Grid (WLCG) e-Infrastructure, powered by the
Particle Physics scientific community, built up to satisfy the computing needs
of processing, storing and analysing all the data coming from the Large Hadron
Collider particle accelerator.

Years of major developments followed, mainly focused on providing the new
functionalities, hindering the adoption of software engineering methodologies.
The software delivered was by then little tested, resulting in constant
rollbacks and patches applied to production systems. Thereby, the stability and
reliability of the e-Infrastructures was deeply compromised, affecting
negatively their exploitation by the scientific communities.

Soon it become apparent that a balance was needed between new features addition
and the reliability of the software produced. Thus, a substantial part of the
software lifecycle should be devoted to deal with the quality of software being
delivered. At this point, evolving software engineering practices were
considered and gradually adopted to define the quality procedures, team
organization and pilot testbeds.

>>>>

\IEEEPARstart{W}hen e-Infrastructures in Europe were first conceived, new
services to access distributed resources -- data, computing, instrumental --
were initially considered. Major investments were then injected in order to
develop distributed solutions that provided a seamless and transparent access
to those resources, available for the researchers. One of the most known examples is Worldwide
Large Hadron Collider Computing Grid (WLCG) global computing infrastructure, powered by the
Particle Physics scientific community, built up to satisfy the computing needs
of processing, storing and analysing the huge quantity of data coming from the Large Hadron
Collider particle accelerator.

Years of major developments followed, mainly focused on providing new
functionalities, hindering the adoption of software engineering methodologies.
The software delivered was by then less tested, resulting in constant
rollbacks and patches applied to production systems. Thereby, the stability and
reliability of the e-Infrastructures was often compromised, affecting
negatively their exploitation by the scientific communities.

Soon it become apparent that a balance was needed between new features addition
and the reliability of the software produced. As a consequence, a substantial part of the
software lifecycle would have to be devoted to deal with the quality of software being
delivered. At this point, evolving software engineering practices were
considered and gradually adopted to define the quality procedures, team
organization and pilot testbeds.
